### 2014-11-11 locks.tex
Change N Queens section.

@@

Table~\ref{queue_instructions} presents the total queue instructions for several
programs. All the coordinated programs require far more queue instructions than
the regular version. For instance, the coordinated SSSP program requires 3 to 4
times more instructions to perform queue operations. Interestingly, the buffered
version requires even more instructions although it performs less overall
coordination. It is likely that this is due to increased fact computation and
subsequent queue manipulation. We argue that the 5\% performance improvement in the
buffered version is due mostly to reduced locking.

For the HT program, the coordinated version requires twice more queue
instructions than the regular version. As expected, the local-only version
requires less queue operations because the number of derived facts is reduced.

\begin{topfig}
\footnotesize\begin{tabular}{ | l || c | c | c | c | c |}
\hline
\multirow{2}{*}{\textbf{Program}} & \multicolumn{5}{c |}{\textbf{Queue
   Instructions}} \\ \cline{2-6}
& 1 & 2 & 4 & 8 & 16 \\ \hline
\hline
SSSP - Regular & 298K & 305K & 309K & 286K & 325K \\ \hline
SSSP - Coordinated & 961K & 919K & 903K & 900K & 922K \\ \hline
SSSP - Buffered & 939K & 961K & 999K & 926K & 1M \\ \hline
\hline
HT - Buffered & 38M & 49M & 50M & 52M & 54M \\ \hline
HT - Coordinated & 57M & 63M & 95M & 89M & 82M \\ \hline
HT - Local-Only & 57M & 63M & 91M & 80M & 74M \\ \hline
\hline
LBP - Regular & 23M & 26M & 24M & 17M & 12M \\ \hline
SBP - Coordinated & 95M & 93M & 88M & 81M & 75M \\ \hline
\hline
MiniMax - Regular & 24M & 26M & 27M & 26M & 27M \\ \hline
MiniMax - Coordinated & 51M & 51M & 52M & 51M & 52M \\ \hline
\end{tabular}
\scap{queue_instructions}{Number of queue \scare{instructions} per program. An
   operation to a regular queue counts as 1 instruction, while an operation to
   a priority queue costs the number of
   \emph{percolate-up}/\emph{percolate-down} steps performed.}
\end{topfig}

>>>>

\paragraph{Queue Instructions}

@@

Table~\ref{queue_instructions} presents the total queue instructions for several
programs. All the coordinated programs require far more queue instructions than
the regular version. For instance, the coordinated SSSP program requires 4 to 5
times more instructions to perform queue operations. Interestingly, the buffered
version requires even more instructions although it performs less overall
coordination. It is likely that this is due to increased fact computation and
subsequent queue manipulation because coordination operations are buffered.
We argue that the 5\% performance improvement in the
buffered version is due mostly to reduced locking.

For the HT program, the coordinated version requires twice more queue
instructions than the regular version. As expected, the local-only version
requires less queue operations because the number of derived facts is reduced.
The MiniMax program also sees twices as many queue instructions using
coordination. This does not happen with the N Queens program, where the number
of queue instructions is far greater when using coordination, especially when
using a small number of threads. We argue that this happens because higher
priority nodes at the bottom are often added to the priority queue, resulting in
many \emph{percolate-up} operations.

\begin{topfig}
\footnotesize\begin{tabular}{ | l || c | c | c | c | c |}
\hline
\multirow{2}{*}{\textbf{Program}} & \multicolumn{5}{c |}{\textbf{Queue
   Instructions}} \\ \cline{2-6}
& 1 & 2 & 4 & 8 & 16 \\ \hline
\hline
SSSP - Regular & 199K & 201K & 205K & 196K & 207K \\ \hline
SSSP - Coordinated & 964K & 938K & 870K & 886K & 963K \\ \hline
SSSP - Buffered & 1M & 1M & 1M & 1M & 1M \\ \hline
\hline
HT - Buffered & 25M & 32M & 34M & 34M & 36M \\ \hline
HT - Coordinated & 51M & 55M & 85M & 80M & 72M \\ \hline
HT - Local-Only & 51M & 55M & 79M & 74M & 68M \\ \hline
\hline
LBP - Regular & 15M & 17M & 16M & 11M & 8M \\ \hline
SBP - Coordinated & 96M & 97M & 87M & 83M & 77M \\ \hline
\hline
MiniMax - Regular & 16M & 19M & 19M & 19M & 20M \\ \hline
MiniMax - Coordinated & 51M & 51M & 51M & 51M & 52M \\ \hline
\hline
N Queens - Regular & 8K & 15K & 16K & 17K & 20K \\ \hline
N Queens - Coordinated & 57M & 18M & 676K & 125K & 41K \\ \hline
\end{tabular}
\scap{queue_instructions}{Number of queue \scare{instructions} per program. An
   operation to a regular queue counts as 1 instruction, while an operation to
   a priority queue costs the number of
   \emph{percolate-up}/\emph{percolate-down} steps performed.}
\end{topfig}

### 2014-11-11 nqueens.tex
use new scap and fig macros
Further updates.
Change N Queens section.

Since computation goes from the top row to the bottom row, not all placements of
nodes on threads may be equally performant. This is especially true because the
bottom rows tend to perform the most work. The best placement is then to split the
board vertically, so that, in optimal conditions, each thread gets the same
number of columns.  Our experiments show that, on shared memory, it does not
matter much if we start with a bad placement since node stealing overcomes those
issues by improving load balancing dynamically. But what if we want run the
program on a distributed system, where moving nodes between workers can be
expensive? We measured the performance of the program without node stealing in
order to assess how important data placement would be for such systems.
We used the default partition of the program as it is, which results in the board being
split by rows and another configuration using the coordination action
\texttt{set-cpu(A, vertical(X, Y))} that forces nodes to be moved to another
thread using a vertical partitioning. Note that \texttt{X} and \texttt{Y} are
the coordinates of the node in the chessboard.

Figure~\ref{results:nqueens} shows the performance results of coordinating the
placing of nodes on threads without work stealing. There is a noticiable
performance improvement when using an optimal data placement strategy. Note
however that in the plot we do not take into account the costs of moving
nodes between workers. In a real system, the performance gap may not be as wide
since in the worst case there may be 144 moves.

\begin{figure}[ht!]
   \begin{center}
      \subfloat[]{\includegraphics[width=5cm]{results/13queens.png}}
   \end{center}
   \caption{Experimental results for the 13 Queens program. The coordinated
      version is on average 2 times faster than the version with bad data
      placement.}
   \label{results:nqueens}
\end{figure}

>>>>

Since computation goes from the top row to the bottom row, not all placements of
nodes on threads may be equally performant. This is especially true because the
bottom rows tend to perform the most work. The best placement is then to split the
board vertically, so that, in optimal conditions, each thread gets the same
number of columns. Our experiments show that, on shared memory, it does not
matter much if we start with a bad placement since node stealing overcomes those
issues by improving load balancing dynamically. However, we noticed interesting
cache effects when prioritizing nodes in bottom rows. The N Queens program
builds and shares lists representing valid board states and needs to iterate
over those lists often when checking if a new position does not invalidate a state.
We thus used coordination axioms \texttt{set-default-priority(A, X)} and \texttt{set-cpu(A,
vertical(X, Y)}, where \texttt{X} is the row number (0 is the top row) and
\texttt{Y} is the column number.

We first ran the coordinated program using dynamic partitioning (with work
stealing) and saw some performance improvements. However, using dynamic
partitioning is bad for memory locality since threads will steal nodes often.
We then added \texttt{set-static(A)} as a coordination axiom and saw
performance improvement, especially when the board columns are perfectly
partitioned by the number of threads (in our case 13 threads).
Figure~\ref{results:nqueens} shows the performance gains saw for the 13 Queens
program.

\begin{topfig}
   \begin{center}
      \subfloat[]{\includegraphics[width=5.5cm]{results/13queens.png}}
   \end{center}
   \scap{results:nqueens}{Experimental results for the 13 Queens program.}
\end{topfig}
